Fishery Markets and Large-Scale Marine Conservation

### 2019-05-30 docs/old_ms/Markets_Conservation_May23.tex
Added all files

% Place your abstract within the special {sciabstract} environment.
\begin{sciabstract}
Will rights-based approaches to managing natural resources facilitate, or relegate large-scale conservation? We study this question in the context of large-scale marine protected areas, and ask whether a country who commits to large-scale conservation is destined to suffer commensurate losses in fishery income.  Indeed, this is a question island (and other) nations all over the world face as they consider whether, and how, to comply with various international commitments to massively expand marine protection. We use a combination of spatial bioeconomic modeling, which accounts for fisheries markets such as the vessel-day trading scheme in the Pacific, and vessel-level empirical data on fishing behavior before and after MPA implementation, to assess how market design incentivizes, or relegates conservation. The answer hinges critically on two design features of the rights-based fishery market. First, only if trading of harvest or vessel-days is allowed across countries can a conservation-minded country capture the economic spillover benefits of their conservation.  Instead, if trading is forbidden, countries are likely to suffer significant losses from conservation. Second, as rights are re-allocated between countries periodically, the re-allocation must depend only on biological features, and not on a country's past effort or catch. Otherwise, the reallocation acts as a punishment to the conserving country. Overall, these results show that seemingly innocuous design features of a market can have indelible impacts on a country's willingness to engage in large-scale conservation; these results are confirmed with an empirical case study from the Phoenix Islands Protected Area, one of the world's largest marine reserves.
\end{sciabstract}

@@

Recognizing a need to secure biodiversity, blue carbon, rare habitats, and other ecosystem services, various international bodies have committed to dramatically expand marine protection around the world. Currently only about 3\% of the world's ocean is strongly protected \cite{sala_2018}, and these commitments call for 10\%, 30\%, or even up to 50\% of the ocean to be off-limits to fishing. Fundamentally, to achieve these goals, huge swaths of sovereign nations' waters will have to be closed to fishing. What incentives might motivate a country to engage in such large-scale marine protection? Would any country rationally close 20\%, 50\%, or even 100\% of its waters to fishing if this meant losing all fishing revenue from its borders? We show that a country's incentives for large-scale conservation hinge critically on an unsuspected culprit: The design of fishing rights.

We are motivated by a real-world institution called the Parties to the Nauru Agreement (PNA), which is one of the most successful, profitable, and well-orchestrated international cartels from any sector of the economy. Like an OPEC for tuna, the PNA is a nine-country coalition of Pacific Island states that collectively manages tuna fishing in its members' waters\cite{havice_2013,aqorau_2018}. These waters collectively account for 14.5 million Km\textsuperscript{2} (an area four times larger than the continental United States), and over 60\% of skipjack catches in the Western Central Pacific \cite{havice_2013}.

In addition to highly productive tuna fisheries, the PNA waters provide a wealth of other ecosystem benefits; this explains why international attention has focused there for targeted conservation efforts to help achieve the previously mentioned targets. But the PNA member nations derive enormous benefits from leasing tuna fishing rights to foreign fleets, often exceeding 50\% of a country's revenue. A natural concern arises: Even if the closure creates net benefits to the world, and perhaps even to tuna fisheries Pacific-wide, the economic loss to the conserving country could still be enormous, and may prevent it from engaging in any meaningful conservation. Intuitively, then, it seems that a country's willingness to engage in large-scale conservation will depend crucially on whether it can successfully capture a sufficiently high fraction of the benefits that are derived from that conservation. We show how the rights-based fisheries market can be designed (if it is new) or modified (if it already exists) to ensure that the conserving country is rewarded, not punished, for large-scale conservation.

We develop a 10-patch spatial bioeconomic model of a fishery operating under a vessel-day scheme (VDS), and where conservation is being considered. Patches 1 - 9 operate under a VDS and Patch 10 represents the high seas in open access. Patch 1 considers closing a portion of its waters ($R$, between 0 and 1). There are a total of 45,000 vessel-days, equally distributed between the 9 VDS-managed patches. The fishery revenue to these patches is the product of within-patch vessel-days and vessel-day price. Fishing vessels decide to fish in each patch based on available biomass (\emph{i.e.} biomass not within a reserve), and are willing to pay a vessel-day price that equates to their marginal profits at the last unit of effort applied. Total biomass is equally distributed across the 10 patches. The stock remains within each patch during the fishing season, but growth from escapement is equally distributed across the 10 patches.

We test the implications of conservation under two market designs: one in which effort trading between countries is not allowed, and one where vessel-days can be traded between countries. On both cases we solve for the equilibrium vessel-day price and effort redistribution. We quantify the losses by comparing revenues from each scenario to a benchmark scenario without conservation. The fishery is simulated for a range of reserve sizes and assumptions about within-patch stock movement. We then examine the role of vessel-day allocation rules, where periodic allocation is a weighted combination of historical within-patch effort and biomass. Finally, we use vessel-tracking data to corroborate our results by looking at the Phoenix Islands Protected Area, implemented in 2015 by Kiribati within the PNA. We use the same data and insights from our model to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary, targeting an 80\% closure of its Exclusive Economic Zone \cite{cimino_2019}.

We begin by simulating a fishery where effort is not traded between countries. We find that a spatial closure in Patch 1 will always result in a loss in revenues (Fig. \ref{fig:PNA_model}A). Closing a greater proportion of the patch results in greater costs, but these decrease if within-patch stock movement ($\theta$) increases. The closure-to-cost ratio for any reserve size is greater than 1:1 when stock movement is low (\emph{i.e.} $\theta < 0.2$), implying that a 30\% closure would result in at least a 30\% loss in revenues. Even for a highly mobile stock, a reserve reduces the amount of biomass that is available for harvest in Patch 1, which reduces a vessel's willingness to pay to fish in such waters (Fig. \ref{fig:vessel_day_price_no_trading_plot}). When countries can not trade, the costs of conservation are large and incurred by Patch 1, but the benefits are received by the eight other patches.

How would a different market design, where trading between countries is allowed, change these results? To answer this we simulate the same fishery, but now allow for vessel-days to be traded between patches. A closure in Patch 1 lowers the value of vessel-days in that patch. Vessel-days in patches 2 - 9 are worth more causing vessel-days from Patch 1 to be sold to Patches 2 - 9. Trading continues until vessel-day prices are the same across all patches (Fig. \ref{fig:vessel_day_price_with_trading_plot}). Under this market design, we find that revenue losses are less than 1\%, compared to the benchmark scenario with no reserve (Fig. \ref{fig:PNA_model}). This means that between 88\% and 99\% of the costs can be avoided if markets are designed to enable trading (Fig. \ref{fig:PNA_model}C).

A second concern about spatial closures in fishing effort markets is that of the re-allocation rules and their implications for costs and long-term sustainability of the reserves. If re-allocation depends on the historical patch-level effort, trading vessel-days to other countries would imply a reduction in allocations through time. We simulate the fishery 50 years into the future and annually re-allocate vessel-days based on a 7-year running mean of patch-level effort and biomass. We use a range of allocation rules to weight effort ($\alpha$) and biomass ($1 - \alpha$) and compare the revenues to a fishery operating for 50 years without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in losses of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, these costs are reduced as the weight on historical biomass is increased. A biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in costs as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation.

We have shown that rights-based fisheries management may reduce the costs of conservation if trading is allowed, and that the costs can be further reduced if the allocation of rights is based on biomass, rather than on effort. We now move to an empirical setting where we use the implementation of one of the world's largest MPAs (implemented in the PNA) and vessel-tracking data to identify the effort displacement and possible costs, as well as the behavioral responses of fishing vessels. We then combine our predictions and empirical results to provide likely outcomes of upcoming large-scale MPAs in the region.

Kiribati implemented the Phoenix Islands Protected Area in January, 2015 (Fig. \ref{fig:PNA_map}). With an extension of 397,447 Km\textsuperscript{2}, this 11.5\% closure effectively displaced all fishing effort within its boundaries \cite{mccauley_2016,mcdermott_2018}. We use identification of fishing activity via Automatic Identification Systems (AIS) provided by Global Fishing Watch to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018 \cite{kroodsma_2018}. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation, and we refer to them as the ``displaced vessels''. The remaining 28 vessels never fished in PIPA waters, and we refer to them as ``non-displaced vessels''. The group with the remaining 221 vessels contains vessels that were not continuously observed before and after the implementation of PIPA, and we refer to these as ``other vessels''.

We use the tracking data to calculate the time that vessels spent within PNA waters (Figs. \ref{fig:all_PS_VDS_year}A-B). Our observations match the 45,000 vessel-days allowed at the PNA level, and roughly match the 11,000 vessel-days allocated to Kiribati \cite{yeeting2018stabilising}. We find that displaced vessels fished 922 less days (a 10\% reduction) in Kiribati between 2014 and 2015, immediately after PIPA was implemented. For the same period, these vessels spent just 48 days less in PNA waters, suggesting that trading might have facilitated relocation of effort within PNA waters.

On the other hand, non-displaced and other vessels spent 1,621 and 2,034 additional days in Kiribati in 2015, relative to 2014. The same pattern is observed at the PNA level, with vessels spending an additional 3,789 and 3,691 vessel-days, respectively. These changes result in a net decrease of fishing effort within Kiribati, from 12,282 in 2014 to just 7,542 in 2018, with displaced vessels driving the decrease (Figs. \ref{fig:all_PS_VDS_year}A). The spatial redistribution patterns of displaced vessels relative to non-displaced vessels suggest that some relocated to other waters in Kiribati (\emph{i.e.} Gilbert islands and Line islands), but also the Marshal Islands and the high seas (Fig \ref{fig:fishing_raster_diff}).

Our analysis suggests that the implementation of PIPA resulted in a decrease in fishing effort applied to Kiribati. But does this result in losses to Kiribati? We do not explicitly observe trading between countries, but combining different sources of information we are able to paint a comprehensive picture. We use country-level fisheries revenue data reported by the Pacific Islands Forum Fisheries Agency (FFA) for the 2008 - 2016 period to complement our analysis (Fig. \ref{fig:revenues}A). We observe that Kiribati's reported revenue increased from \$127.3 million USD in 2014 to \$148.8 million USD in 2015, before decreasing to \$118.3 million USD in 2016. This increase and subsequent decrease in revenues match the patterns we observe for Kiribati in 2014 - 2016 (Figs. \ref{fig:all_PS_VDS_year}A), and observed effort was a good predictor of revenue (Fig. \ref{fig:revenue_FFA_effort_GFW}). Unless At the PNA-level, total revenues showed a net increase of \$64.7 and \$28 million USD for 2015 and 2016, respectively (Fig. \ref{fig:total_PNA_revenues}).

We also inspect the crowding effects that may arise by applying more (or the same) fishing effort over less fishing area by inspecting the spatial overlap between displaced and non-displaced vessels.These groups interacted more with each other after the implementation of PIPA (Table \ref{tab:sp_corr}, Fig. \ref{fig:sp_corr}). The number of cells with presence from both fleets and spatial correlation between fleets increase by a factor of four and three, respectively. Environmental variation may certainly drive this, but the NINO4 anomaly alone explained just 3\% and 7\% of the variation in our crowding measures, compared to the 70\% variation explained when accounting for PIPA implementaiton. We observe similar patterns when repplicating the crowding exercise for Kiribati's EEZ only (Table \ref{tab:KIR_sp_corr}, Fig \ref{fig:KIR_sp_corr}).

@@

On October 28, 2015, the President of Palau signed into law the Palau National Marine Sanctuary (PNMS) Act. Starting in December 2020, this Act will close 500,000 Km\textsuperscript{2} (about 80\% of Palau’s EEZ) to commercial fishing activities, creating the 14th largest protected area in the world and effectively displacing all effort within its waters (Fig. \ref{fig:PNA_map}). As a PNA member, Palau receives 700 purse seine vessel-days each year. Their longline fishery is also managed under a vessel-day scheme, with 10,500 vessel-days per year (Fig. \ref{fig:PLW_GFW_ts}). Unlike the purse seine vessel-days, longline vessel-days are not tradable across countries and are worth much less, at about \$200 USD each.

Analysis of vessel-tracking data revealed that longliners and purse seiners spent 56.2\% and 91.7\% of their time within PNMS (Fig. \ref{fig:plw_2018}). Our predictions and empirical analyses of PIPA suggest that Palau may be able to reduce the costs if purse seine vessel-days are traded with other PNA members or it continues to sell them to the United States under the South Pacific Tuna Treaty (also known as the Multilateral Treaty on Fisheries; \$12,500/day). However, the current reallocation rule for purse seine vessel-days is based on historical effort and biomass, suggesting that the allocation would decrease in time.

International goals over the next decade have set ambitious targets for marine conservation. Plans to implement large-scale Marine Protected Areas must consider the opportunity costs that come with such interventions \cite{smith_2010}, and identify sustainable financing mechanisms that compensate losses and incentivize long-term marine conservation \cite{mallin_2019}. We show that by combining rights-based fisheries management, a market design that allows trading, and a biomass-based allocation rule, large-scale marine conservation can be incentivized by significantly reducing costs. The Parties to the Nauru Agreement have shown that rights-based management of transboundary resources can result in large management and economic benefits \cite{havice_2013,aqorau_2018}. By facilitating trade and modifying the re-allocation rule, they may become important leaders in effective marine conservation.

@@

% PNA model output for costs
\begin{figure}[htbp]
\centering
\includegraphics{img/PNA_model.pdf}
\caption{\label{fig:PNA_model}Cost of spatial closures in a vessel-day fishery (vertical axis) as a function of reserve size ($R$; horizontal axis) and movement ($\theta$; line color). Costs are shown for Patch 1 when there is no trading (A) and when trading is allowed (B). Costs avoided by tradig are shown in (C).}
\end{figure}

@@

% Palau Raster and PNMS
\begin{figure}
\centering
\includegraphics{img/plw_2018.png}
\caption{\label{fig:plw_2018}Longline and purse seine fishing effort in Palaw during 2018 at a 0.5 degree resolution. The red polygon shows the Palau National Marine Sanctuary, containing 56\% and 91\% of longline and purse seine fishing effort, respectively.}
\end{figure}

>>>>

% Place your abstract within the special {sciabstract} environment.
\begin{sciabstract}
Will rights-based approaches to managing natural resources facilitate, or impede large-scale conservation? We study this question in the context of large-scale marine protected areas, and ask whether a country that commits to large-scale conservation is destined to suffer commensurate losses in fishery income.  Indeed, this is a question maritime nations all over the world face as they consider whether, and how, to comply with various international commitments to expand marine protection. We use a combination of spatial bioeconomic modeling, which accounts for fisheries markets such as the vessel-day trading scheme in the Pacific Ocean, and vessel-level empirical data on fishing behavior before and after MPA implementation, to assess how market design incentivizes, or penalizes conservation. The answer hinges critically on two design features of the rights-based fishery market. First, only if trading of harvest or vessel-days is allowed across countries can a conservation-minded country capture the economic spillover benefits of their conservation actions.  Instead, if trading is forbidden, countries are likely to suffer losses from conservation. Second, when rights are allocated between countries, the allocation must depend only on biological features, and not on economic factors such as a country's past effort or catch. Otherwise, the allocation process acts as a punishment to the conserving country. Overall, these results show that seemingly innocuous design features of a market can have indelible impacts on a country's willingness to engage in large-scale conservation; these results are confirmed with an empirical case study from the Phoenix Islands Protected Area, one of the world's largest marine reserves. With the right design features, large-scale protected areas such as the Phoenix Islands can be achieved at close to zero cost to the implementing country.
\end{sciabstract}

@@

Recognizing a need to secure biodiversity, blue carbon, rare habitats, and other ecosystem services, various international bodies have committed to dramatically expand marine protection around the world. Currently only about 3\% of the world's ocean is strongly protected \cite{sala_2018}, and these commitments call for 10\%, 30\%, or even up to 50\% of the ocean to be off-limits to fishing. Fundamentally, to achieve these goals, huge swaths of sovereign nations' waters will have to be closed to fishing. What incentives might motivate a country to engage in such large-scale marine protection? Would any country rationally close 20\%, 50\%, or even 100\% of its waters to fishing if this meant losing all fishing revenue from within its borders? We show that a country's incentives for large-scale conservation hinge critically on an unsuspected culprit: the design of fishing rights.

We are motivated by a real-world institution called the Parties to the Nauru Agreement (PNA). Like an OPEC for tuna, the PNA is a nine-country coalition of Pacific Island states that collectively manages tuna fishing in its members' waters\cite{havice_2013,aqorau_2018}. These waters collectively account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental United States), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}.

In addition to highly productive tuna fisheries, the PNA waters provide a wealth of other ecosystem benefits, hence the focus on large-scale conservation efforts in the region. The PNA member nations derive enormous benefits from leasing tuna fishing rights to foreign fleets, in some cases exceeding 50\% of a country's annual revenue. A natural concern arises: even if the closure creates net benefits to the world, and perhaps even to tuna fisheries Pacific-wide, the economic loss to the conserving country could still be extremely large, and may prevent it from engaging in any meaningful conservation. Intuitively, then, it seems that a country's willingness to engage in large-scale conservation will depend crucially on whether it can successfully capture a sufficiently high fraction of the benefits that are derived from that conservation. We show how a rights-based fisheries market can be designed (if it is new) or modified (if it already exists) to ensure that the conserving country is rewarded, not punished, for large-scale conservation.

We develop a 10-patch spatial bioeconomic model of a fishery operating under a vessel-day scheme (VDS), and where conservation is being considered. A VDS is a type of rights-based management approach used by the PNA where each country is given an allocation of vessel-days that can then be sold to fishing firms. A vessel-day for country $i$ gives a firm the right to have one vessel fish in country $i$'s EEZ for 24 hours. Patches 1 - 9 operate under a VDS and Patch 10 represents the high seas in open access. Patch 1 considers closing a portion of its waters as a reserve ($R$, between 0 and 1). There are a total of 45,000 vessel-days, equally distributed between the 9 VDS-managed patches. The fishery revenue to each of these patches is the product of within-patch vessel-days and vessel-day price. Fishing vessels decide to fish in each patch based on available biomass (\emph{i.e.} biomass not within a reserve), and are willing to pay a vessel-day price that equates to the marginal profits of the last unit of effort applied. Total biomass is equally distributed across the 10 patches. The stock remains within each patch during the fishing season, but growth from escapement is equally distributed across the 10 patches.

We test the implications of conservation under two market designs: one in which effort trading between countries is not allowed, and one where vessel-days can be traded between countries. In both cases we solve for the equilibrium vessel-day price and effort redistribution. We quantify the losses by comparing revenues from each scenario to a benchmark scenario without any conservation action. The fishery is simulated for a range of reserve sizes and assumptions about within-patch stock movement. The second market design feature we investigate is the allocation of rights. We vary the degree to which periodic allocation is based on historical within-patch effort versus the actual abundance of fish biomass in a patch. Finally, we use vessel-tracking data and official reports of government revenue to corroborate our results by looking at the Phoenix Islands Protected Area, implemented in 2015 by Kiribati within the PNA. \textbf{We use the same data and insights from our model to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary, currently targeting an 80\% closure of its Exclusive Economic Zone \cite{cimino_2019}.}

We begin by simulating a fishery where effort is not traded between countries. We find that a spatial closure in Patch 1 will always result in a loss in revenues to Patch 1 (Fig. \ref{fig:PNA_model}A). Closing a greater proportion of the patch results in greater costs, but these decrease if within-patch stock movement ($\theta$) increases, allowing vessels to harvest the stock within the remaining open area. The closure-to-cost ratio for any reserve size is greater than 1:1 when stock movement is low (\emph{i.e.} $\theta < 0.2$), implying that a 30\% closure would result in at least a 30\% loss in revenues. Even for a highly mobile stock, a reserve reduces the amount of biomass that is available for harvest in Patch 1, which reduces a vessel's willingness to pay to fish in such waters (Fig. \ref{fig:vessel_day_price_no_trading_plot}). When countries cannot trade, the costs of conservation are large and incurred by Patch 1, but the benefits are received by the eight other patches (Figure S2 shows that the increase in revenue ranges from 0\% to 6\%).

How would a different market design, where trading between countries is allowed, change these results? To answer this we simulate the exact same fishery, but now allow for vessel-days to be traded across patches. A closure in Patch 1 lowers the value of vessel-days in that patch. Vessel-days in patches 2 - 9 are now worth more causing vessel-days from Patch 1 to be traded to Patches 2 - 9. Trading continues until vessel-day prices are the same across all patches (Fig. \ref{fig:vessel_day_price_with_trading_plot}). Under this market design, we find that revenue losses are less than 1\%, compared to the benchmark scenario with no reserve (Fig. \ref{fig:PNA_model}). This means that between 88\% and 99\% of the costs of conservation can be avoided if markets are designed to enable trading (Fig. \ref{fig:PNA_model}C).

A second concern about spatial closures in fishing effort markets is that of the allocation rules and their implications for costs and the long-term sustainability of the closures. If allocation rules updates based on the historical patch-level effort, then trading vessel-days to other countries would imply a future reduction in allocation. We simulate the fishery 50 years into the future and annually re-allocate vessel-days based on a 7-year running mean of patch-level effort and biomass. We use a range of allocation rules to weight effort ($\alpha$) and biomass ($1 - \alpha$) and compare the revenues to a fishery operating for 50 years without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in losses of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, these costs are reduced as the weight on historical biomass is increased. A biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in costs as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation.

We have shown that rights-based fisheries management may reduce the costs of conservation if trading is allowed, and that the costs can be further reduced if the allocation of rights is based on biomass, rather than on effort. We now move to an empirical setting where we use the implementation of one of the world's largest MPAs and vessel-tracking data to identify the effort displacement and possible costs of conservation. \textbf{We then combine our predictions and empirical results to predict the likely outcomes of upcoming large-scale MPAs in the region.}

Kiribati implemented the Phoenix Islands Protected Area in January 2015 (Fig. \ref{fig:PNA_map}). With an extension of 397,447 Km\textsuperscript{2}, this 11.5\% closure effectively displaced all fishing effort within its boundaries \cite{mccauley_2016,mcdermott_2018}. We use identification of fishing activity via Automatic Identification Systems (AIS) provided by Global Fishing Watch to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018 \cite{kroodsma_2018}. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation, and we refer to them as the ``displaced vessels''. The remaining 28 vessels never fished in PIPA waters, and we refer to them as ``non-displaced vessels''. The group with the remaining 221 vessels contains vessels that were not continuously observed before and after the implementation of PIPA, and we refer to these as ``other vessels''.

We use the tracking data to calculate the time that vessels spent within PNA waters (Figs. \ref{fig:all_PS_VDS_year}A-B). Our observations match the 45,000 vessel-days allowed at the PNA level, and roughly match the 11,000 vessel-days historically allocated to Kiribati \cite{yeeting2018stabilising}. We find that displaced vessels fished 922 less days (a 10\% reduction) in Kiribati between 2014 and 2015, immediately after PIPA was implemented. For the same period, these vessels spent just 48 days less in PNA waters, suggesting that trading might have facilitated relocation of effort within PNA waters.

On the other hand, non-displaced and other vessels spent 1,621 and 2,034 additional days in Kiribati in 2015, relative to 2014. The same pattern is observed at the PNA level, with vessels spending an additional 3,789 and 3,691 vessel-days, respectively. These changes result in a net decrease of fishing effort within Kiribati, from 12,282 in 2014 to just 7,542 in 2018, with displaced vessels driving the decrease (Figs. \ref{fig:all_PS_VDS_year}A). The spatial redistribution patterns of displaced vessels relative to non-displaced vessels suggest that some relocated to other waters in Kiribati (\emph{i.e.} the Gilbert islands and Line islands), but also the Marshall Islands and the high seas (Fig \ref{fig:fishing_raster_diff}).

Our analysis suggests that the implementation of PIPA resulted in a decrease in fishing effort within Kiribati's waters. But does this result in losses to Kiribati? We do not explicitly observe trading between countries, but combining different sources of information we are able to paint a comprehensive picture. We use country-level fisheries revenue data reported by the Pacific Islands Forum Fisheries Agency (FFA) for the 2008 - 2016 period to complement our analysis (Fig. \ref{fig:revenues}A). We observe that Kiribati's reported revenue increased from \$127.3 million USD in 2014 to \$148.8 million USD in 2015, before decreasing to \$118.3 million USD in 2016. This increase and subsequent decrease in revenues matches the AIS patterns we observe for Kiribati in 2014 - 2016 (Fig. \ref{fig:all_PS_VDS_year}A), and observed AIS effort is a reliable predictor of government fishing license revenue across all PNA countries from 2008 to 2016 (Fig. \ref{fig:revenue_FFA_effort_GFW}). At the PNA level, total revenues showed a net increase of \$64.7 and \$28 million USD for 2015 and 2016 respectively (Fig. \ref{fig:total_PNA_revenues}).

We also inspect the crowding effects that may arise by applying more (or the same) fishing effort over less fishing area by inspecting the spatial overlap between displaced and non-displaced vessels. These groups interacted more with each other after the implementation of PIPA (Table \ref{tab:sp_corr}, Fig. \ref{fig:sp_corr}). The number of cells with presence from both fleets and the spatial correlation between fleets increases by a factor of four and three, respectively. Environmental variation may certainly drive this, but the NINO4 anomaly alone explained just 3\% and 7\% of the variation in our crowding measures, compared to the 70\% variation explained when accounting for PIPA implementaiton. We observe similar patterns when replicating the crowding exercise for Kiribati's EEZ only (Table \ref{tab:KIR_sp_corr}, Fig \ref{fig:KIR_sp_corr}).

@@

\textbf{On October 28, 2015, the President of Palau signed into law the Palau National Marine Sanctuary (PNMS) Act. Starting in December 2020, this Act will close 500,000 km\textsuperscript{2} (about 80\% of Palau’s EEZ) to commercial fishing activities, creating the 14th largest protected area in the world (Fig. \ref{fig:PNA_map}). As a PNA member, Palau receives 700 purse seine vessel-days each year. Their longline fishery is also managed under a vessel-day scheme, with 10,500 vessel-days per year (Fig. \ref{fig:PLW_GFW_ts}). Unlike the purse seine vessel-days, longline vessel-days are currently not tradable across countries and are worth much less, at about \$200 USD each.}

\textbf{Analysis of vessel-tracking data revealed that longliners and purse seiners spent 56.2\% and 91.7\% of their time within PNMS (Fig. \ref{fig:plw_2018}). Our predictions and empirical analyses of PIPA suggest that Palau may be able to reduce the cost of lost license revenue if purse seine vessel-days are traded with other PNA members or it continues to sell them to the United States under the South Pacific Tuna Treaty (also known as the Multilateral Treaty on Fisheries; \$12,500/day). However, the current reallocation rule for purse seine vessel-days is based on historical effort and biomass, suggesting that the allocation would decrease over time. Further, the fact that longline vessel days are currently not tradable means that these losses will be larger than would be the case with trading.}

International goals over the next decade have set ambitious targets for marine conservation. Plans to implement large-scale Marine Protected Areas must consider the opportunity costs that come with such interventions \cite{smith_2010}, and identify sustainable financing mechanisms that compensate losses and incentivize long-term marine conservation \cite{mallin_2019}. We show that by combining rights-based fisheries management, a market design that allows trading, and a biomass-based allocation rule, large-scale marine conservation can be incentivized by eliminating the costs. The Parties to the Nauru Agreement have shown that rights-based management of transboundary resources can result in large management and economic benefits \cite{havice_2013,aqorau_2018}. By facilitating trade and modifying the re-allocation rule, they may become pioneers in effective large-scale marine conservation.

@@

% PNA model output for costs
\begin{figure}[htbp]
\centering
\includegraphics{img/PNA_model.pdf}
\caption{\label{fig:PNA_model}Cost of spatial closures in a vessel-day fishery (vertical axis) as a function of reserve size ($R$; horizontal axis) and movement ($\theta$; line color). Costs are shown for Patch 1 when there is no trading (A) and when trading is allowed (B). Costs avoided by trading are shown in (C). Dashed red line is a 45 degree line.}
\end{figure}

@@

% Palau Raster and PNMS
\begin{figure}
\centering
\includegraphics{img/plw_2018.png}
\caption{\label{fig:plw_2018}Longline and purse seine fishing effort in Palau during 2018 at a 0.5 degree resolution. The red polygon shows the Palau National Marine Sanctuary, containing 56\% and 91\% of longline and purse seine fishing effort, respectively.}
\end{figure}
